\myTitle

### 2024-05-13 abstract.tex
Eric final edits

%!TEX main.tex

While a best practice for evaluating the behavior of genetic clustering algorithms 
on empirical data is to conduct parallel analyses on simulated data, these types 
of simulation techniques often involve sampling genetic data with replacement. 
In this paper we demonstrate that sampling with replacement, especially with large 
marker sets, inflates the perceived statistical power to correctly assign individuals 
back to source populations---a phenomenon we refer to as resampling-induced, 
spurious power inflation (RISPI). To address this issue, we present \gscramble{}, a 
simulation approach in R for creating biologically informed individual genotypes from 
empirical data that: 1) samples alleles from populations \textit{without} replacement, 
2) segregates alleles based on species-specific recombination rates. \gscramble{} is 
a flexible simulation approach that allows users to create pedigrees of varying complexity 
in order to simulate admixed genotypes and it allows users to track haplotype blocks 
from the source populations through the pedigrees. We demonstrate the functionality of 
\gscramble{} with both simulated and empirical data sets and highlight additional uses of 
the package that users may find valuable.

>>>>

%!TEX root = main.tex

While a best practice for evaluating the behavior of genetic clustering algorithms 
on empirical data is to conduct parallel analyses on simulated data, these types 
of simulation techniques often involve sampling genetic data with replacement. 
In this paper we demonstrate that sampling with replacement, especially with large 
marker sets, inflates the perceived statistical power to correctly assign individuals
(or the alleles that they carry)
back to source populations---a phenomenon we refer to as resampling-induced, 
spurious power inflation (RISPI). To address this issue, we present \gscramble{}, a 
simulation approach in R for creating biologically informed individual genotypes from 
empirical data that: 1) samples alleles from populations \textit{without} replacement, 
2) segregates alleles based on species-specific recombination rates. This framework
makes it possible to simulate admixed individuals in a way that respects the physical
linkage between markers on the same chromosome and which does not suffer
from RISPI.  This is achieved in \gscramble{} by allowing 
users to specify pedigrees of varying complexity 
in order to simulate admixed genotypes, segregating and tracking haplotype blocks 
from different source populations through those pedigrees, and then sampling---using
a variety of permutation schemes---alleles from empirical data into those haplotype blocks.
We demonstrate the functionality of 
\gscramble{} with both simulated and empirical data sets and highlight additional uses of 
the package that users may find valuable.

### 2024-05-13 body-text-of-paper.tex
Eric final edits

In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals of
mixed ancestry---individuals with proportions of their genome attributed to multiple subpopulations, sensu \citet{pritchard2000inference}.
Various processes could contribute to the appearance of mixed ancestry in a clustering analysis.  Most directly, individuals may represent the
influences of gene flow among populations, occurring in recent or prior generations, with the observed complex ancestry patterns accurately
representing contributions from multiple sampled populations.  However, similar patterns may result from functionally different patterns of
population structuring.  For example, such patterns might be expected when clinal systems governed by isolation by distance are described
as discrete genetic clusters.
Alternatively, individuals of mixed ancestry could reflect immigration from outlying populations (e.g., exogenous immigration), with true source
populations not sufficiently
represented in the sample to be among the characterized $K$ subpopulations.
Furthermore, the characterization of admixed individuals may reflect a limitation of the statistical power of the genetic data,
as a function of both the marker set and sample size, to resolve the true underlying patterns of genetic structure.
 The challenge for researchers then becomes correctly identifying the ecological and/or statistical processes by which observed complex
 ancestry patterns were created.

@@

In response to these needs, we developed \gscramble{} – a simulation approach that samples alleles from respective populations without
replacement, thus maintaining genetic differences between the clusters commensurate to those in the empirical dataset, and eliminating the effect of RISPI.
Further, by simulating individual genotypes based on user-defined pedigrees of arbitrary complexity, \gscramble{} allows for the simulation of admixed genotypes resulting from any conceivable category of (non-inbred) hybrid, while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
By integrating species-specific recombination rates, \gscramble{} simulates biologically informed genotypes that mirror empirical data.
We develop and illustrate \gscramble{} through the use of single nucleotide polymorphism (SNP) genotypes, though it is equally applicable to data sets with hundreds or thousands of
microhaplotypes \citep{baetscher2018microhaplotypes} or microsatellites
\citep{zhan2017megasat}.

@@

Because the simulations were done in a na\"{i}ve way that effectively assumes that the population
allele frequencies are identical to those observed in the samples themselves, we hypothesized that ADMIXTURE would return estimates of
$q_A$ that were
centered around the simulated values, even though samples $A$ and $B$ were drawn from the same
population. In other words, we expected to observe RISPI.
We further expected that ADMIXTURE would return more accurate estimates of the simulated
$q_A$ values when $N$ was smaller and the number of loci was larger - 
attributes that magnify the effect of RISPI.
We summarized the results with boxplots of the ADMIXTURE-inferred $q_A$ values,
 for the 480 newly simulated individuals (and for the original samples, $A$ and $B$, in the
 case of unsupervised clustering),
for each combination of $L$, $N$, and $n$.

@@

Genetic material is segregated through
a GSP using these steps, done upon each node in an order such that the steps
are run on each node's parents before being run on the node itself:
\begin{enumerate}
\item Each genome in a triangle above a founder node represents a gamete
carrying a single, haploid copy of the genome.
These gametes are united within the founder individual and then
recombination occurs between each chromosome
with probability $\frac{1}{2}$ and within each chromosome according to
a user-specified recombination map to create two new gametes, which are
recombined versions of the original ones.  These two gametes are then segregated at random down the edge (or edges) from the founder node. Upon each edge, $i$ from founder node $f$, $G^-_{f,i}$ gametes are segregated. 
\item Each non-founder node, $x$, will have exactly two edges coming
into it---one from each parent.  Each of the $G^+_{x,1}$ gametes coming in from the
first edge is united with a single one of the $G^+_{x,2}$ gametes coming
in on the second edge (in a random order).
\item If a non-founder node $x$ has an edge to a sample node,
a randomly chosen $S_x$ of the united pairs of gametes are delivered
to the sample node to constitute the $S_x$ samples taken from individual
node $x$. Any remaining united pairs of gametes are treated as in (4) below.
\item United pairs of gametes (including those remaining after delivery
to a sample node) in each non-founder node $x$ are segregated into
gametes with recombination, and each of these gametes is assigned, in
random order, to the gamete edges proceeding downward out of the node,
with the number of gametes assigned to the $i\thh$ downward edge being
$G^-_{x,i}$.
\end{enumerate}
At the end of this process, the samples simulated (as united pairs of gametes)
into the sample nodes will contain all of the genetic material (and no more) found among the founders, but that
material will have been segregated and recombined according to the pedigree.  Thus the
samples obtained from a GSP can be simulated with the expected admixture fractions
for any hybrid category that can be specified by a pedigree.  They have also been simulated
in a way that 1) reflects physical linkage---each individual inherits segments of the genome
simulated using a recombination map; 2) does not duplicate any genetic material---the sampling of
genetic material is entirely without replacement so it will not induce RISPI; and 3) no genetic material has been lost---all the genetic material of the founders
is represented in the samples, maximizing the number of simulated individuals and markers available to
use in downstream analyses.

@@

We illustrate the use of \gscramble{} in three example cases. The first is
a simple simulation to show that, unlike sampling with replacement, the
approach of \gscramble{} does not produce RISPI.
The next two are examples using empirical data sets.
The first of these involves
simulating a small number ($<\,$70) of markers that are fixed, or nearly
fixed, for alternate alleles between two species of trout: rainbow trout ({\em 
Oncorhynchus mykiss}) and cutthroat trout ({\em O. clarkii}). The goal of the 
simulation study is to determine how accurately individuals of different hybrid
categories between the species can be identified, and whether the use of linked 
markers deflates the estimates of uncertainty to an unacceptable degree.  The 
second example involves many more markers ($\approx$14,000) and 
investigates the accuracy for
identifying individuals admixed between closely related groups of feral
pigs in Missouri.

@@

For this example, we repeat the simulation described in {\em \nameref{sec:rispi-sim}};
however, after simulating individuals in the two samples drawn from the same population,
we create data sets of admixed individuals using \gscramble{}, rather than sampling
gene copies with replacement.  As before, individuals were simulated with fractions of
ancestry from sample $A$ in $\{0, \frac{1}{8}, \frac{1}{4}, \frac{3}{8}, \frac{1}{2}, \frac{5}{8}, \frac{3}{4}, \frac{7}{8}, 1\}$, but these were simulated from pedigrees using gscramble.  We limited
this simulation to one of the scenarios that showed extreme RISPI when sampling gene
copies with replacement: $L=1,000$ markers with sizes of samples $A$ and $B$ being
$N=50$. This simulation was repeated 240 times, and the results of each replicate were analyzed with
ADMIXTURE to estimate the ancestry fractions of each simulated individual.

\subsubsection*{Rainbow/cutthroat trout hybrids in California}

We used the archived data from \citet{rizza2023limited}, drawing upon their 65 markers
that are fixed or are nearly fixed for alternate alleles between rainbow trout and cutthroat trout.
They designated 634 fish as pure cutthroat trout and 213 as pure steelhead.  We used the
genotypes of these individuals as input to \gscramble{}.

@@

In the original paper, \citet{rizza2023limited} identified 15 $F_1$s,
1 $F_2$,
6 cutthroat backcrosses ($BC_1$s) and 
6 second-generation cutthroat backcrosses ($BC_2$s),
and 1 fish of an unknown hybrid category that may have been the
product of a $BC_2$ crossed with a $BC_1$.   In our simulations, we sought to simulate
a similar collection of hybrid individuals, so, for each replicate, we used
\gscramble{} to simulate 16 $F_1$s, 4 $F_2$s, 6 $BC_1$s, 4 $BC_2$s, 4 $BC_3$s,  4 individuals
that were the products of a $BC_1$ mating with a $BC_2$, and finally, 4 individuals that were simulated pure cutthroat. Individuals not used as founders
in \gscramble{} were retained as pure cutthroat and pure steelhead.  The goal of the simulation
was to assess how well these different hybrid categories can be distinguished from one another
when using either all 65 markers, or just the 30 markers occurring on different chromosome arms
used by  \citet{rizza2023limited}.

@@

\subsubsection*{Rainbow/cutthroat trout hybrids in California}

The barplots in Fig.~\ref{fig:newhyb-sims-barplot} show the distribution of posterior
probabilities of the different hybrid categories. 
%%%%%%%%%%
\begin{figure}
\newcommand{\nhbpcap}{\footnotesize Posterior probabilities for simulated
individuals of different trout hybrid categories.  Each pair of plots shows the results
for individuals of the true hybrid category listed atop the pair  (see
 Table~\protect\ref{tab:newhybcats} for the meanings of the names).  The top panel of each pair
shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated with the original 30 markers,
plus an additional  35 (65 total), some of which are linked
on the same chromosome. Each individual appears as a thin vertical line, posterior probabilities of
each inferred hybrid category are indicated by extent of different colors on those lines.  Individuals
are sorted according to maximum {\em a posteriori} hybrid category, and, within that by the
maximum posterior probability. }
{\centering
\includegraphics[width=0.98\columnwidth]{figures/newhybs-texed-crop.pdf}
}
\caption[\nhbpcap]{\nhbpcap}
\label{fig:newhyb-sims-barplot}
\end{figure}
%%%%%%%%%%
These are similar to STRUCTURE or ADMIXTURE barplots;
however, rather than plotting the ancestry fraction, the different colors are showing the
posterior probabilities of different hybrid categories.  Each individual is represented by a thin
vertical line, and the individuals have been arranged from left to right and the colors have been
arranged from top to bottom throughout different sections of the plot to make it easy to see
the patterns.  The {\em maximum-a-posteriori} (MAP) hybrid category for any individual is represented
by the color that is on the bottom of the bar, so it is easy to see the shifts where the MAP hybrid
category changes from one to another.  It is apparent over all individuals simulated to be
either CCT (cut-throat trout), $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
hybrid categories were correct for largely the same number of individuals whether 30 or 65 markers
were used. By contrast, for the individuals simulated to be of the complex hybrid category
$BC_1$--CCT~$\times$~$BC_2$--CCT, the MAP estimate from 65 markers was correct more often 
than from 
30 markers.  However, it is also apparent that using 65 markers, some of which are physically
linked, in the software {\em NewHybrids}---which assumes that all markers are unlinked---yields 
estimates of uncertainty that are inaccurate.  Tellingly, when the estimates using 65 markers are 
incorrect (for example, when $BC_1$--CCT individuals have a maximum posterior of being in the 
$BC_2$--CCT category,
or when individuals that are truly of the $BC_1$--CCT~$\times$~$BC_2$--CCT category have a
maximum posterior of being from the $BC_2$--CCT category), many of those incorrect posteriors
are quite high ($>0.90$).  In fact, using 65 markers yields more incorrect {\em and high} posteriors 
than does using the 30 unlinked markers. So, as expected, ignoring physical linkage (e.g., as 
implemented in {\sc NewHybrids}) 
can yield a poor estimate of the uncertainty around different inferences.

@@

The simulated hybrid classes produce
the expected range of admixture with $F_1$ and $F_2$ individuals having mean $Q=0.5$, with $F_2$ individuals having
more variation (Figure~\ref{fig:pig-sim-qs}A).
Backcrossed individuals have higher ancestry associated with the backcrossed population with mean $Q$
approximately $0.75$ ($\BC_1$) and $0.87$ ($\BC_2$) (Figure~\ref{fig:pig-sim-qs}A). Interestingly, the distribution of the $Q$ values for $F_1$'s varies more across the different genetic clusters than it does for the other
hybrid categories.
When the estimated $Q$ values for the $F_1$ hybrids are separated by the population combinations,
$F_1$ hybrids for Pop1-Pop2 have the least variation around $q=0.5$ (standard deviation
= 0.008), while Pop1-Pop3 and Pop2-Pop3
have higher variation (standard deviation = 0.011-0.016) (Figure~\ref{fig:pig-sim-qs}B). Also, all $Q$
distributions for $F_1$ hybrids, regardless of population combination, center around
$0.50$ except for the two distributions
of Pop3, which have a mean of $Q=0.49$ (Figure~\ref{fig:pig-sim-qs}B).
The differences among these
distributions of $Q$ values for the simulated data reflect the genetic differences observed
in the empirical data, specifically, that genetic differentiation is highest between Pop1 and Pop2 and that
Pop3 is the most admixed population with ancestry contributions from both Pop1 and Pop2.
The results emphasize difficulties that may arise in estimating admixture fractions between
groups that have experienced historical gene flow.

@@

Here, we present \gscramble{}, a simulation approach for creating biologically informed individual 
genotypes of different hybrid classes. Two integral features of \gscramble{} are: 
1) alleles are sampled from populations without replacement and 
2) alleles are segregated based on species-specific recombination rates. 
Our results show that both 
of these features are especially important for accurately reflecting empirical data from large genetic 
marker sets (thousands of markers), 
which are common from  both custom SNP-chip genotyping and next generation sequencing methods. Sampling alleles without 
replacement, and thereby preserving the genetic differences (or lack thereof) of populations is 
essential for avoiding RISPI---an issue that is exacerbated as the number of 
alleles sampled for an individual increases. Additionally, we demonstrate the importance of accounting for 
the physical linkage of markers when large marker sets are used and that treating physically linked loci as 
independent erroneously reduces the variation in the admixture 
fractions of simulated individuals.

\gscramble{} is conducive for analysis of numerous types of marker sets available from contemporary data generation 
techniques. As mentioned above, 
one key feature of \gscramble{} is the integration of recombination rates, thereby accounting for physical 
linkage which is expected among larger contemporary data sets. Furthermore, \gscramble{} performs marker 
segregation with computational efficiency---for the four different GSPs used with the invasive wild pig data set, 
the mean run time of each simulation with 14,233 SNP loci was 2.3 seconds (range: 1.9-5.7 $s$). 
While \gscramble{} is equally applicable to data sets with 
microhaplotypes \citep{baetscher2018microhaplotypes} or microsatellites \citep{zhan2017megasat}, 
it is not suited for genotype likelihood data (i.e. from low-coverage whole genome sequencing). 
Specifically, simulations conducted within \gscramble{} involve the segregation of gametes within the 
specified pedigrees and, therefore, requires genotypes to be resolved.

We envision \gscramble{} to be relevant for a diverse range of potential applications. In our 
provided empirical examples with trout and wild pig data, we demonstrate the basics of 
creating hybrid classes. In much the same manner, \gscramble{} has been used to evaluate the statistical power 
of hybrid classification used to differentiate wild boar hybrids from domestic pigs 
\citep{smyser2024probgen}. Specifically \citet{smyser2024probgen} used the segregation options in \gscramble{} 
 to preserve founder individuals' genotypes within populations in contrast to \textit{a priori} shuffling alleles 
within populations. This retains the subpopulation structure present within certain populations---thereby, better reflecting the 
actual admixture variation expected from wild boar and domestic pig hybrids. We think \gscramble{} 
may also be useful for assessing relatedness estimates using pedigrees of varying complexity. Here, it 
is important to note that each time a gamete is segregated in a GSP, its complement is also 
segregated---an example of antithetical variate sampling in Monte Carlo methods. Thus, in the 
example of relatedness, fixed specified founders should be repeatedly "mated" with one offspring 
sampled per iteration to avoid sampling of anticorrelated genotypes.

\textbf{Eric's notes}: 
Somewhere here I want to mention that the Anderson et al. (2008) approach for simulating
mixtures of genotypes to assess how well one can estimate mixture proportions---the fraction
of individuals from each of a set of baseline or reference populations---allows the user to
simulate, if desired, more genotypes than occur in the reference samples.  This is because
a trick is done in the inference stage to remove the correlation between the simulated individual
and the reference sample, relative to the true population, by removing the simulated individual's
genotypes from its own reference population when calculating the likelihood that its genotype
originated from each of the reference populations.  To do the same when doing simulations for
assessing power using ADMIXTURE or STRUCTURE would require hacking the underlying code
of those programs, which is not desirable.  So, for these cases we have used the sampling
without replacement approach.

Also, we need to discuss the fact that, because each time a gamete is segregated in the
pedigree, its complement is also segregated, this is an example of antithetical variate sampling
in Monte Carlo.

\section*{Acknowledgements}
Research was supported in part by the USDA-APHIS-National Feral Swine Damage
Management Program. We are grateful for the numerous Wildlife Services and Missouri
Department of Conservation employees that contributed to genetic sampling of wild pigs
within southeastern Missouri. We thank Yniv Palti for providing flanking sequences and the linkage map for 57K markers for making a recombination map for the cutthroat/rainbow trout example. Writing of this paper was facilitated by the Mobile High Altitude Venue for Ecological Analysis,
Genetics and Statistics, through which ECA, MGD, and TJS were scientists-in-residence on location in
Cheyenne, WY for 3 days in October 2023. This is contribution number mHAVEAGAS-003.
This work utilized the Alpine high performance computing resource at the University of Colorado Boulder.
Alpine is jointly funded by the University of Colorado Boulder, the University of Colorado Anschutz, Colorado State University,
and the National Science Foundation (award 2201538).

>>>>

In using clustering algorithms to describe genetic structure among natural populations, researchers frequently identify individuals of
mixed ancestry---individuals with proportions of their genome attributed to multiple subpopulations, sensu \citet{pritchard2000inference}.
Various processes could contribute to the appearance of mixed ancestry in a clustering analysis.  Most directly, individuals may represent the
influences of gene flow among populations, occurring in recent or prior generations, with the observed complex ancestry patterns accurately
representing contributions from multiple sampled populations.  However, similar patterns may result from functionally different patterns of
population structuring.  For example, such patterns might be expected when clinal systems governed by isolation by distance are described
as discrete genetic clusters.
Alternatively, individuals of mixed ancestry could reflect immigration from outlying populations (e.g., exogenous immigration), with true source
populations not sufficiently
represented in the sample to be among the characterized $K$ subpopulations.
Furthermore, the characterization of admixed individuals may reflect a limitation of the statistical power of the genetic data (as a function of both the marker set and sample size) to resolve the true underlying patterns of genetic structure.
 The challenge for researchers then becomes correctly identifying the ecological and/or statistical processes by which observed complex
 ancestry patterns were created.

@@

In response to these needs, we developed \gscramble{}---a simulation approach that samples alleles from respective populations without
replacement, thus maintaining genetic differences between the clusters commensurate to those in the empirical dataset, and eliminating the effect of RISPI.
Further, by simulating individual genotypes based on user-defined pedigrees of arbitrary complexity, \gscramble{} allows for the simulation of admixed genotypes resulting from any conceivable category of (non-inbred) hybrid, while allowing the tracking of haplotype blocks from source populations within the specified pedigree.
By integrating species-specific recombination rates, \gscramble{} simulates biologically informed genotypes that mirror empirical data.
We develop and illustrate \gscramble{} through the use of single nucleotide polymorphism (SNP) genotypes, though it is equally applicable to data sets with hundreds or thousands of
microhaplotypes \citep{baetscher2018microhaplotypes} or microsatellites
\citep{zhan2017megasat}.

@@

Because the simulations were done in a na\"{i}ve way that effectively assumes that the population
allele frequencies are identical to those observed in the samples themselves, we hypothesized that ADMIXTURE would return estimates of
$q_A$ that were
centered around the simulated values, even though samples $A$ and $B$ were drawn from the same
population. In other words, we expected to observe RISPI.
We further expected that ADMIXTURE would return estimates of $q_A$ that were closer
to the simulated values when $N$ was smaller and the number of loci was larger---attributes that magnify the effect of RISPI.
We summarized the results with boxplots of the ADMIXTURE-inferred $q_A$ values,
 for the 480 newly simulated individuals (and for the original samples, $A$ and $B$, in the
 case of unsupervised clustering),
for each combination of $L$, $N$, and $n$.

@@

Genetic material is segregated through
a GSP using these steps, done upon each node in an order such that the steps
are run on each node's parents before being run on the node itself:
\begin{enumerate}
\item Each genome in a triangle above a founder node represents a gamete
carrying a single, haploid copy of the genome.
These gametes are united within the founder individual and then
recombination occurs between each chromosome
with probability $\frac{1}{2}$ and within each chromosome according to
a user-specified recombination map to create two new gametes, which are
recombined versions of the original ones.  These two gametes are then segregated at random down the edge (or edges) from the founder node. Upon each edge, $i$ from founder node $f$,  a number, $G^-_{f,i}$, of gametes are segregated. 
\item Each non-founder node, $x$, will have exactly two edges coming
into it---one from each parent.  Each of the $G^+_{x,1}$ gametes coming in from the
first edge is united with a single one of the $G^+_{x,2}$ gametes coming
in on the second edge (in a random order).
\item If a non-founder node $x$ has an edge to a sample node,
a randomly chosen $S_x$ of the united pairs of gametes are delivered
to the sample node to constitute the $S_x$ samples taken from individual
node $x$. Any remaining united pairs of gametes are treated as in (4) below.
\item United pairs of gametes (including those remaining after delivery
to a sample node) in each non-founder node $x$ are segregated into
gametes with recombination, and each of these gametes is assigned, in
random order, to the gamete edges proceeding downward out of the node,
with the number of gametes assigned to the $i\thh$ downward edge being
$G^-_{x,i}$.
\end{enumerate}
At the end of this process, the samples simulated (as united pairs of gametes)
into the sample nodes will contain all of the genetic material (and no more) found among the founders, but that
material will have been segregated and recombined according to the pedigree.  Thus the
samples obtained from a GSP can be simulated with the expected admixture fractions
for any hybrid category that can be specified by a pedigree.  They have also been simulated
in a way that 1) reflects physical linkage---each individual inherits segments of the genome
simulated using a recombination map; 2) does not duplicate any genetic material---the sampling of
genetic material is entirely without replacement so it will not induce RISPI; and 3) no genetic material has been lost---all the genetic material of the founders
is represented in the samples, maximizing the number of simulated individuals and markers available to
use in downstream analyses.

@@

We illustrate the use of \gscramble{} in three example cases. The first is
a simple simulation to show that, unlike sampling with replacement, the
approach of \gscramble{} does not produce RISPI.
The next two are examples using empirical data sets.
The first of these involves
simulating a small number ($<\,$70) of markers that are fixed, or nearly
fixed, for alternate alleles between two species of trout: steelhead trout ({\em 
Oncorhynchus mykiss}) and cutthroat trout ({\em O. clarkii}). The goal of the 
simulation study is to determine how accurately individuals of different hybrid
categories between the species can be identified, and whether the use of linked 
markers deflates the estimates of uncertainty to an unacceptable degree.  The 
second example involves many more markers ($\approx$14,000) and 
investigates the accuracy for
identifying individuals admixed between closely related groups of feral
pigs in Missouri.

@@

For this example, we repeat the simulation described in {\em \nameref{sec:rispi-sim}};
however, after simulating individuals in the two samples drawn from the same population,
we create data sets of admixed individuals using \gscramble{}, rather than sampling
gene copies with replacement.  As before, individuals were simulated with fractions of
ancestry from sample $A$ in $\{0, \frac{1}{8}, \frac{1}{4}, \frac{3}{8}, \frac{1}{2}, \frac{5}{8}, \frac{3}{4}, \frac{7}{8}, 1\}$, but these were simulated from pedigrees using \gscramble{}.  We limited
this simulation to one of the scenarios that showed extreme RISPI when sampling gene
copies with replacement: $L=1,000$ markers with sizes of samples $A$ and $B$ being
$N=50$. This simulation was repeated 240 times, and the results of each replicate were analyzed with
ADMIXTURE to estimate the ancestry fractions of each simulated individual.

\subsubsection*{Steelhead/cutthroat trout hybrids in California}

We used the archived data from \citet{rizza2023limited}, drawing upon their 65 markers
that are fixed or are nearly fixed for alternate alleles between steelhead trout and cutthroat trout.
They designated 634 fish as pure cutthroat trout and 213 as pure steelhead.  We used the
genotypes of these individuals as input to \gscramble{}.

@@

In the original paper, \citet{rizza2023limited} identified 15 $F_1$s,
1 $F_2$,
6 cutthroat backcrosses ($BC_1$--CCTs) and 
6 second-generation cutthroat backcrosses ($BC_2$--CCTs),
and 1 fish of an unknown hybrid category that may have been the
product of a $BC_2$ crossed with a $BC_1$.   In our simulations, we sought to simulate
a similar collection of hybrid individuals, so, for each replicate, we used
\gscramble{} to simulate 16 $F_1$s, 4 $F_2$s, 6 $BC_1$s, 4 $BC_2$s, 4 $BC_3$s,  4 individuals
that were the products of a $BC_1$ mating with a $BC_2$, and finally, 4 individuals that were simulated pure cutthroat. Individuals not used as founders
in \gscramble{} were retained as pure cutthroat and pure steelhead.  The goal of the simulation
was to assess how well these different hybrid categories can be distinguished from one another
when using either all 65 markers, or just the 30 markers occurring on different chromosome arms
used by  \citet{rizza2023limited}.

@@

\subsubsection*{Steehlead/cutthroat trout hybrids in California}

The barplots in Fig.~\ref{fig:newhyb-sims-barplot} show the distribution of posterior
probabilities of the different hybrid categories. 
%%%%%%%%%%
\begin{figure}
\newcommand{\nhbpcap}{\footnotesize Posterior probabilities for simulated
individuals of different trout hybrid categories.  Each pair of plots shows the results
for individuals of the true hybrid category listed atop the pair  (see
 Table~\protect\ref{tab:newhybcats} for the meanings of the names).  The top panel of each pair
shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated with the original 30 markers,
plus an additional  35 (65 total), some of which are linked
on the same chromosome. Each individual appears as a thin vertical line, posterior probabilities of
each inferred hybrid category are indicated by extent of different colors on those lines.  Individuals
are sorted according to maximum {\em a posteriori} hybrid category, and, within that by the
maximum posterior probability. }
{\centering
\includegraphics[width=0.98\columnwidth]{figures/newhybs-texed-crop.pdf}
}
\caption[\nhbpcap]{\nhbpcap}
\label{fig:newhyb-sims-barplot}
\end{figure}
%%%%%%%%%%
These are similar to STRUCTURE or ADMIXTURE barplots;
however, rather than plotting the ancestry fraction, the different colors are showing the
posterior probabilities of different hybrid categories.  Each individual is represented by a thin
vertical line, and the individuals have been arranged from left to right and the colors have been
arranged from top to bottom throughout different sections of the plot to make it easy to see
the patterns.  The {\em maximum-a-posteriori} (MAP) hybrid category for any individual is represented
by the color that is on the bottom of the bar, so it is easy to see the shifts where the MAP hybrid
category changes from one to another.  It is apparent over all individuals simulated to be
either CCT (cutthroat trout), $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
hybrid categories were correct for largely the same number of individuals whether 30 or 65 markers
were used. By contrast, for the individuals simulated to be of the complex hybrid category
$BC_1$--CCT~$\times$~$BC_2$--CCT, the MAP estimate from 65 markers was correct more often 
than from 
30 markers.  However, it is also apparent that using 65 markers, some of which are physically
linked, in the software {\em NewHybrids}---which assumes that all markers are unlinked---yields 
estimates of uncertainty that are inaccurate.  Tellingly, when the estimates using 65 markers are 
incorrect (for example, when $BC_1$--CCT individuals have a maximum posterior of being in the 
$BC_2$--CCT category,
or when individuals that are truly of the $BC_1$--CCT~$\times$~$BC_2$--CCT category have a
maximum posterior of being from the $BC_2$--CCT category), many of those incorrect posteriors
are quite high ($>0.90$).  In fact, using 65 markers yields more incorrect {\em and high} posteriors 
than does using the 30 unlinked markers. So, as expected, ignoring physical linkage (e.g., as 
implemented in {\sc NewHybrids}) 
can yield a poor estimate of the uncertainty around different inferences.

@@

The simulated hybrid classes produce
the expected range of admixture with $F_1$ and $F_2$ individuals having mean $Q=0.5$, and $F_2$ individuals having
more variation (Figure~\ref{fig:pig-sim-qs}A).
Backcrossed individuals have higher ancestry associated with the backcrossed population with mean $Q$
approximately $0.75$ ($\BC_1$) and $0.87$ ($\BC_2$) (Figure~\ref{fig:pig-sim-qs}A). Interestingly, the distribution of the $Q$ values for $F_1$ hybrids varies more across the different genetic clusters than it does for the other
hybrid categories.
When the estimated $Q$ values for the $F_1$ hybrids are separated by the population combinations,
$F_1$ hybrids for Pop1-Pop2 have the least variation around $q=0.5$ (standard deviation
= 0.008), while Pop1-Pop3 and Pop2-Pop3
have higher variation (standard deviation = 0.011-0.016) (Figure~\ref{fig:pig-sim-qs}B). Also, all $Q$
distributions for $F_1$ hybrids, regardless of population combination, center around
$0.50$ except for the two distributions
of Pop3, which have a mean of $Q=0.49$ (Figure~\ref{fig:pig-sim-qs}B).
The differences among these
distributions of $Q$ values for the simulated data reflect the genetic differences observed
in the empirical data, specifically, that genetic differentiation is highest between Pop1 and Pop2 and that
Pop3 is the most admixed population with ancestry contributions from both Pop1 and Pop2.
The results emphasize difficulties that may arise in estimating admixture fractions between
groups that have experienced historical gene flow.

@@

Here, we present \gscramble{}, an R package implementing a simulation
approach for creating biologically informed individual 
genotypes of different hybrid classes. Two integral features of \gscramble{} are: 
1) alleles are sampled from populations without replacement and 
2) alleles are segregated based on species-specific recombination rates. 
Our results show that both 
of these features are especially important for accurately reflecting empirical data from large genetic 
marker sets (thousands of markers), 
which are common from both custom SNP-chip genotyping and next generation sequencing methods. Sampling alleles without 
replacement, and thereby preserving the genetic differences (or lack thereof) of populations is 
essential for avoiding RISPI---an issue that is exacerbated as the number of 
genetic markers increases. Additionally, we demonstrate the importance of accounting for 
the physical linkage of markers when large marker sets are used, and that treating physically linked loci as 
independent erroneously reduces the variation in the admixture 
fractions of simulated individuals.

\gscramble{} is suitable for analysis of numerous types of marker sets available from contemporary data generation 
techniques. As mentioned above, 
one key feature of \gscramble{} is the integration of recombination rates, thereby accounting for physical 
linkage which is expected among high-density markers sets. Furthermore, \gscramble{} performs marker 
segregation with computational efficiency---for the four different GSPs used with the invasive wild pig data set, 
the mean run time of each simulation with 14,233 SNP loci was 2.3 seconds (range: 1.9-5.7 $s$). 
While \gscramble{} is equally applicable to data sets with 
microhaplotypes \citep{baetscher2018microhaplotypes} or microsatellites \citep{zhan2017megasat}, 
it is not built to use genotype likelihoods (i.e., from low-coverage whole genome sequencing). 
Specifically, simulations conducted within \gscramble{} involve the segregation of gametes within the 
specified pedigrees and, therefore, require called genotypes (though missing data is acceptable).

We envision \gscramble{} to be relevant for a diverse range of potential applications. In our 
provided empirical examples with trout and wild pig data, we demonstrate the basics of 
creating hybrid classes. In much the same manner, \gscramble{} has been used to evaluate the statistical power 
of hybrid classification used to differentiate wild boar hybrids from domestic pigs 
\citep{smyser2024probgen}. Specifically \citet{smyser2024probgen} used the segregation options in \gscramble{} 
 to preserve founder individuals' genotypes in contrast to \textit{a priori} shuffling alleles 
within populations. This retains the subpopulation structure present within certain populations---thereby, better reflecting the 
actual admixture variation expected from wild boar and domestic pig hybrids. We think \gscramble{} 
may also be useful for assessing relatedness estimates using pedigrees of varying complexity. Here, it 
is important to note that each time a gamete is segregated in a GSP, its complement is also 
segregated---an example of antithetical variate sampling in Monte Carlo methods \citep{hammersley}.
Thus, in the  example of relatedness, fixed specified founders should be repeatedly "mated" with one offspring 
sampled per iteration to avoid sampling of anticorrelated genotypes.

Here, we have demonstrated the consequences of failing to consider physical linkage among loci 
given the availability of a recombination map. However, in the absence of a recombination map, 
researchers with a basic understanding of the genome size and the number of chromosomes in 
the study organism could still benefit from using \gscramble{} to simulate genetic data under
an assumed, approximate linkage map. To do 
so, users of \gscramble{} could randomly distribute markers across, and within, the 
known number of chromosomes to produce a conceptual marker map (*.map within the PLINK 
file structure), and assume basic rates of recombination (e.g., 1 cM per 1 Mb). The resulting 
simulations from \gscramble{} would produce admixture fractions that capture the expected 
variation from known linkage patterns, in contrast to the reduced variation produced from 
assuming independence (Figure 5). Characterizing uncertainty in this way would provide caution for 
researchers to not overinterpret conclusions drawn in the absence of a recombination map.

\section*{Acknowledgements}
Research was supported in part by the USDA-APHIS-National Feral Swine Damage
Management Program. We are grateful for the numerous Wildlife Services and Missouri
Department of Conservation employees that contributed to genetic sampling of wild pigs
within southeastern Missouri. We thank Yniv Palti for providing flanking sequences and the linkage map for 57K markers for making a recombination map for the steelhead/cutthroat trout example. Writing of this paper was facilitated by the Mobile High Altitude Venue for Ecological Analysis,
Genetics and Statistics, through which ECA, MGD, and TJS were scientists-in-residence on location near
Cheyenne, WY for 3 days in October 2023. This is contribution number mHAVEAGAS-003.
This work utilized the Alpine high performance computing resource at the University of Colorado Boulder.
Alpine is jointly funded by the University of Colorado Boulder, the University of Colorado Anschutz, Colorado State University,
and the National Science Foundation (award 2201538).
